%% Abstract (Resumo em inglês)

The society has reached technology to create autonomous vehicles and to connect different devices and machinery to each other in order to exchange data and to optimize production efficiency.  With this technology, soon it will be possible to achieve better methods to guide blind and visually impaired (BVI) users in their daily activities. The products currently available in the market have a number of limitations and do not satisfy BVI users. We believe that one of the reasons behind this problem is that BVI are not members of the development team or are not consulted by these. 

The lack of an efficient solution for BVI users navigation became even larger with the SARS-CoV2 pandemic, in which people had to avoid contact with each other and not touch other surfaces.

The purpose of this paper is to use Virtual Reality (VR) to test and evaluate different design of BVI products and to verify if non-BVI users have the same mental demand and situation awareness as a BVI user when using assistive products. The idea is to use the VR as a testing ground where a BVI user can try different assistive solutions in different scenarios. By doing so, the user becomes part of the product design and evaluation, resulting in better and more user-friendly products. The proposed method includes not only the setup of the virtual environment but also the use of physiological sensors and subjective tests in order to assess the mental workload and situational awareness in different situations and/or using different versions of product under development.

To illustrate the proposed method, we use as case study the BVI navigation in a medical clinic submitted to COVID-19 protocols. This case study is chosen due to the current undergoing SARS-CoV-2 pandemic and the impact for BVI people.

The scene was made using Unity3D, a widely used development platform for virtual reality applications. The VR device was the Tobii Eye Tracking VR, a head-mounted display for virtual reality developed using the HTC VIVE. The VR device is used for defining the position and orientation of the user in a virtual environment in Unity. Based on the current situation in the virtual environment, inputs are provided to the user using aural commands and haptics devices. In order to assess the mental workload, physiological sensors, from TEA Captiv T-Sens, are used. Among them, are an electrocardiogram sensor (ECG), to gather heart-rate and heart-rate variance data using, and a galvanic skin reaction sensor (GSR) to collect skin conductance. Beside these sensors, the users are also expected to answer NASA-TLX mental workload assessment test and situation awareness questionnaires.

Among the expected benefits of the proposed method is the flexibility and agility to create different scenarios, as well as the possibility to test all of them in the same physical room. This could not only speed the design of new solutions but also improve the overall quality of the products and to verify the need of a BVI user in the development team of a assistive product.