This chapter discusses a set of selected works from the literature that are related to different aspects of this work. Their selection was performed in the Scopus and Web of Science databases, using keywords such as “human factors”, “virtual reality” and “blindness”. From an initial set of 344 papers, a set of seven were selected as more relevant to this work and are detailed in the next sections.

\section{Virtual reality for BVI users}
\label{sec:vr_without_vision}
\input{Revisao/VR Without Vision/vr_without_vision__siu_et_al}

\section{Feeling of presence in virtual reality}
\label{sec:emotion_presence_vr}
\input{Revisao/Emotion Presence/emotion_presence_vr}

\section{Information for BVI navigation}
\label{sec:bradley_dunlop}
\input{Revisao/Bradley Dunlop/bradley_dunlop}

\section{Audio navigation for BVI}
\label{sec:auditory_navigation}
\input{Revisao/Auditory navigation for BVI/auditory_navigation}

\section{Comparison of assistive devices}
\label{sec:evaluation_spatial_display}
\input{Revisao/Evaluation Spatial Display/evaluation_spatial_display}

\section{Virtual reality in the design process}
\label{sec:vr_cabin}
\input{Revisao/VR Cabin/vr_cabin}

\section{Final Remarks}
\label{sec:final_remarks3}

In this chapter, 7 papers were reviewed with the current thesis in mind. These 7 papers are related to one of the three keywords, or a combination of them: "human factors", "virtual reality" and "blindness".

These papers pointed that BVI users are sensible to sensorial inputs. \ref{sec:vr_without_vision} showed that they were sensible with the vibration of the cane. This sensibility was based on their past experience with canes made from different materials and with vibration devices, but this paper was more about designing a navigation tool, not a framework for testing human factors.

When designing an virtual environment, it is important to consider the feeling of presence, and in the Section \ref{sec:emotion_presence_vr} presented a paper that concluded that to increase this feeling one should allow the user agency inside the environment. That means that the user will feel more present (will forget that he/she is inside the virtual environment) when he/she is induced to interact with the elements from the virtual environment. The current experiment used real furniture and real actors mixed with virtual sounds to increase the sensation that the user was not inside a laboratory, but inside a medical clinic reception.

The Section \ref{sec:bradley_dunlop} was about to papers from the same authors that studied the differences between the information used by BVI and sighted users during the navigation. They found out that the last group used more text-based information, while the first group used information with more words and from a wider variety of categories. Another conclusion an information that was made for BVI users was more mentally demandful for sighted users, and vice-verse. This conclusion was importante for two of the used methods in the thesis, which used audio information for the navigation.

But when is the perfect moment to announce a audio information for BVI users? The Section \ref{sec:auditory_navigation} showed the the moment that an information is transmitted has affect on the accuracy they reach a goal. This conclusion was made by analysing their navigation task with a NASA-TLX and the results confirmed that this tool can be used for their tasks.

One of the evaluations that this work does is comparing different BVI devices. This evalutation was presented in \ref{sec:evaluation_spatial_display} and the author made a similar comparison. They compared an haptic and sound information in different environments. Their conclusion was that the haptic were less acceptable. This work employs two models of haptic devices and a audio method. Is likely that the conclusion will be comparable.

Finally, the last section of this chapter, Section \ref{sec:vr_cabin}, showed the benefits and disavantages of implementing virtual reality with co-design on a design team. Similarly the ideia of this work is to used virtual reality to allow the users to test new concepts of products and give the designer their feedbacks and impressions.

All these papers contributed to the designing, improvement, and decision making of the current experiment. The next chapter will describe all these steps with more details.

