% CAPÍTULO 2: FUNDAMENTAÇÃO TEÓRICA
%   DEVE SER UM CAPÍTULO ENXUTO (LEAN)
%1. Elabore um parágrafo que introduz o capítulo: Este capítulo apresenta (descreva o objetivo do capítulo...). É constituído de N seções a saber...
%2. Desenvolvimento do capítulo: os conteúdos devem ser somente os necessários para o leitor entender a sua contribuição.
%3. Trabalhos relacionados: a última seção do Capítulo 2 deve apresentar o posicionamento da sua contribuição em relação à literatura. É um detalhamento da Justificativa apresentada na Introdução.
%4. Elabore um parágrafo que conclui o capítulo e introduz o capítulo seguinte.

The proposal of this work combines the concepts of co-design and virtual reality with techniques from the area of Human Factors. 

In order to facilitate the understanding of this dissertation, this chapter introduces these concepts and techniques. It starts with the definition of Human Factors, also known as Ergonomics, and describes mental workload and situation awareness, as well as the corresponding assessment methods that are used in this work. It then presents the concept of extended reality and defines virtual reality. Finally, it introduces the principles of co-design.

\section{Human Factor or Ergonomics}
\label{sec:human_factors}

    \input{Fundamentação/Fatores Humanos/human_factors}

\subsection{Mental Workload (MWL)}
\label{sec:mental_workload}

    \input{Fundamentação/Carga Mental/mental_workload}

\subsection{Situation Awareness (SA)}
\label{sec:situation_awareness}

    \input{Fundamentação/Percepção situacional/situation_awareness}

\section{Extended Reality (XR)}
\label{sec:extended_reality}

    \input{Fundamentação/Realidade Extendida/extended_reality}    

\section{Co-Design}
\label{sec:co_design}

    \input{Fundamentação/Co-Design/co_design}

\section{Final remarks}
\label{sec:final_remarks2}

%Colocar um resumo do capitulo descrevendo também como os tres temas abordados são combinados no seu trabalho. Colocar uma seção de final remarks ao final de cada capitulo, com excecao do primeiro e do ultimo.

These topics are fundamental to the comprehension of the results. The Section \ref{sec:human_factors} explains the importance of the Human Factors to improve the efficiency of the user's handling of the tool or machine. It also showed that many different factors can influence this performance and it is important to know how these factors affect this interaction.

Inside this section two important concepts of Human Factors. One of them is the Mental Workload, presented in Section \ref{sec:mental_workload}, an important concept that is unique between an individual and a device and that can deeply affect their interaction. The second is the Situation Awareness, on Section \ref{sec:situation_awareness}, a concept that explains how the mind perceives and processes the elements around. Both these concepts can deeply affect the device manipulation by an user, specially when it is a BVI user.

To this experiment be sucessul, it would need an isolated environment, without the interferance of unwanted effects, and be made inside a medical clinic, which could become an annoyance for the owners. That is all negotiable, if at the time the experiment was made there was not a pandemic, when hospitals and clinic are not a recommended place to visit. An alternative for this is to design a isolated medical clinic inside a controled environment of a laboratory and a fast way to create and to update this environment is throughout Virtual Reality. This is explained inside Section \ref{sec:extended_reality}.

Finally, Co-design, from Section \ref{sec:co_design}, is the process behind one the motivation of this process. A process that brings closer different designers with different backgrounds and experiences in order to design more efficient. In this case, to bring BVI users closer to a research team that need their experience and opinions about their process of navigation and mobility.

The next chapter will dive in 7 papers that were selected as more related to this experiment, and how they correlate to this one.
