%CAPÍTULO 4: ANÁLISE DOS RESULTADOS E DISCUSSÃO

%1. Elabore um parágrafo que introduz o capítulo: Este capítulo apresenta (descreva o objetivo do capítulo ...). É constituído de N seções a saber...
%2. Caso vc tenha aplicado a sua contribuição (modelo, produto, processo etc.) em um caso (empresa, laboratório, simulação etc.), apresente a descrição e análise dos resultados. Na seção de discussão cabem as análises de cenários What-If ou de sensibilidade. Exemplo: se o parâmetro X aumentar de N para N+1, o resultado poderia mudar de Y para Z?
%3. Elabore um parágrafo que conclui o capítulo e introduz o capítulo seguinte.

The purpose of the experiment discussed in this chapter is to investigate the two research questions proposed for this work:

\begin{itemize}%[leftmargin = 3.5em, label = $H_\arabic*$:]
    \item Is it possible to evaluate and compare concepts of assistive device from a human factors’ perspective in a virtual environment? What are the main limitations of the use of a virtual reality environment?
    \item Do non-BVI users, when deprived from their vision, evaluate assistive devices in a similar way as BVI users?
\end{itemize}

For this purpose, the experiment described in Section \ref{ch:metodologia} was performed with the following groups:

\begin{itemize}
    \item Blind group: composed of 4 participants with age varying from 26 to 56, all male, three of them graduated and one with graduation in course. 

    \item Sighted group: composed of 4 participants with age varying from 22 to 31, three males and one woman, all of them graduated.
\end{itemize}

In order to answer the two research questions, this chapter is organized in the following way. Section \ref{sec:results_obj_1} is dedicated to the first question and brings an analysis performed only with data from blind participants. Then, Section \ref{sec:results_obj_2} repeats the same analysis now with data from sighted participants and compare the results with those obtained from blind participants, in order to answer the second research question.

In both sections, the data analysis follows the following sequence:

\begin{itemize}
    \item Analysis of subjective questionnaires:
    \begin{itemize}
        \item NASA-TLX: it aims at assessing the workload perceived by the user in six dimensions, including ‘mental demand’. It is expected that the mental workload would decrease from the ‘first’ to the ‘return’ round. It is also expected that some guidance methods would differ regarding the required mental workload.
        \item Adapted SAGAT: it aims at assessing the situation awareness and the mental map of the user. It is expected that the SAGAT score would increase from the ‘first’ to the ‘return’ round. It is also expected that some guidance methods would differ regarding the required situation awareness provided to the user.
        \item Guidance method’s questionnaire: it aims at assessing the user experience with each method. It is also expected that some guidance methods would differ regarding the score received in this questionnaire.
    \end{itemize}
    \item Analysis of physiological sensors:
    \begin{itemize}
        \item ECG: it aims at assessing the user workload. Two features are extracted from the ECG signal, heartrate (BPM) and heartrate variance (SDNN). It is expected that the heartrate slight decreases from the ‘first’ to the ‘return’ round, while the heartrate variance is expected to slight increase.
        \item GSR: it aims at assessing the user workload and stress. It is expected that the GSR average would increase at every ‘first’ round and then a slight decrease in the ‘return’ round.
    \end{itemize}
\end{itemize}

Particularly in the case of this work, an additional round of tests is added to the experiment in order to investigate the differences between the evaluation performed by BVI users and sighted (non-BVI) users. For this purpose, the experiment is repeated with a set of non-BVI users and the same data is collected. The purpose is to investigate whether or not performing the analysis with non-BVI users could lead to different conclusions.


\section{Evaluation of assistive device from a human factors’ perspective in a virtual environment}
\label{sec:results_obj_1}

\subsection{Subjective data}
\input{Resultados/Nasa/nasa1}
\input{Resultados/Sagat/sagat1}
\input{Resultados/Questionario/questionarios1}

\subsection{Physiological data}

During the experiment, data from two physiological sensors were captured: ECG and GSR. As commonly found in the literature, these data are used to assess mental workload. The corresponding analysis is presented in this section.

\input{Resultados/ECG/ecg1.tex}

\input{Resultados/GSR/gsr1.tex}

\subsection{Final Remarks}

Summarizing the conclusion obtained from the analysis of the data from blind participants, the audio method showed a lower score both for NASA-TLX mental demand and NASA-TLX global score, while the methods that include vibration achieved higher scores. This probably happened because the participants are already used to use sound to guide themselves, especially environmental sounds. The environment sounds used in the scenes were always the same (telephone ringing, laptop keyboard sounds, exterior noise, door opening and closing). It is likely that the participants felt more relaxed when they only had to focus on the sounds around him/her. This is reinforced by the fact that, during experiment with the “audio” method, half of the participants did not ask for any information or the audio command option used only a few times.

The fact that the haptic devices caused a higher workload is probably due to the fact that the users had to learn and get used with them. Besides, for being just conceptual, their precision was not as good as they were expecting. That explains why their results were not as good as the “Base” or “Audio” methods. The NASA-TLX results are correctly related to the satisfaction questionnaires, which scored them as the unsatisfied devices.

As expected, most of the variables from subjective questionnaires (NASA-TLX and SAGAT) show some influence of the round, as expected. On the other hand, the results from the physiological sensors did not show clear tendency. 

The statistical analysis based on ANOVA tests confirmed some of the observations from the bar and box plots. However, in many cases the residual distributions were not homogenous and the statistical analysis was affected by the small number of samples. 

All the blind participants showed a great enthusiasm before, during and after the experiment. They also made a number of recommendations for both the virtual environment and the devices, such as:

\begin{itemize}
    \item The speakers of the HMD are not good enough to give them the precise location of the sound origin
    \item The HMD is too large and cover half of the participant’s face. It gives them a strange sensation, since some of them use the air or the wind feeling on the face to give them hints about the location of walls or other high obstacles;
    \item The precision of the vibration for both haptic belt and virtual cane needs to be improved. It is not enough for them to use the devices. This problem is related to how the HMD set the position of the user in the virtual environment. \\    
    \item The vibration from the haptic belt was not intense enough.
\end{itemize}

\section{Comparison between BVI users and sighted users}
\label{sec:results_obj_2}

This section is dedicated to investigate the second research question of this work: “do non-BVI users, when deprived from their vision, evaluate assistive devices in a similar way as BVI users?”. 

In order to do so, the analysis performed in the previous section is now repeated with the data obtained from sighted participants. However, the data corresponding to the ‘base’ method is omitted, as the daily method used by sighted people is based on their vision.

%\input{Resultados/SubjectiveData2}
\subsection{Subjective data}
\input{Resultados/Nasa/nasa2}
\input{Resultados/Sagat/sagat2}
\input{Resultados/Questionario/questionarios2}

\input{Resultados/PhysiologicalData2} 

\subsection{Final Remarks}

% Mental demand
% • ANOVA - Audio e Mix     VS   não deu resultado
% • Grafico - Não tem uma relação específica VS  Presença de haptico aumenta a demanda mental

% NASA TLX
% • ANOVA - Round  VS  - Efeito do método e do round.
% •• Difere do ANOVA MD     Metodo - 3 grupos
% •••                       Base e Audio menores Carga mental
% ••• Sem padrão no tipo de dispositivo  Haptic alta Carga mental
% 
% SAGAT
% ANOVA - Método tem efeito     VS   Round tem efeito no SAGAT
% Nota menor nas mesmas condições
%

% BPM
% ANOVA - não deu resultado VS IDEM
% Gráfico mostra uma diferença entre os métodos
% Maior BPM que os cegos

% SDNN
% ANOVA - não deu resultado VS IDEM
% Gráfico - "Base" tem uma variância menor
% Valor de SDNN maior que o dos cegos

% GSR
% ANOVA - não deu resultado VS IDEM
% Grafico - métodos muito parecidos VS metodos com impactos diferentes

Differently than the blind users, the results from the mental demand discipline of the NASA-TLX proved that the sight users felt a higher mental demand than the blind users. 

The overall NASA-TLX score also proved a different conclusion than the one in the Section \ref{subsubsec:results_nasa_tlx_1}. For the sighted users, the round impacted more the overall score than the methods, whilst for the blind user were the opposite. This may be because the overall score is composed of 6 dimensions. Probably for the sighted user the mental demand score was higher and for the blind user it was not. But even so, the average score of the sight user was higher than the score of the blind user.

The Adapted SAGAT questionnaire for the sight users proved that the method impacts their situation awareness. The conclusion was different than the one proved by the blind users in the Section \ref{subsubsec:results_adapted_sagat_1}, who felt a bigger impact between the rounds than between the methods. The sight performance was also poorer than the blind user.

The guidance questionnaire of both groups had a similar distribution, both groups enjoy the same methods, "Audio" and "Mixture". The difference was in the last two. The sighted users rather use the "Virtual Cane" and the blind users rather use the "Haptic Belt".

These conclusion show that the sighted users were more sensible to the methods than the blind users, although the effects were different. The blind users were more impacted by the methods than by the rounds, and when impacted by the methods, it was not possible to detect a pattern about the presence or not of a haptic device, as happened with the blind users.

The ECG sensors shown a difference in the heartrate between the methods, but the ANOVA test was not able to prove that difference, the same conclusion of the blind users in the Section \ref{subsubsec:results_ecg_1}. Another observation is that the heartrate frequency of the sighte user was higher than the blind users, meaning that their mental workload was probably higher.

According to the ANOVA test, The heartbeat variance also was not impact by the method or by the rounds, the same conclusion for the blind users in the Section \ref{subsubsec:results_ecg_1}. Graphically there was a small difference in the methods. Despite the results of the heartrate, the variance of the sight user was higher than the results from the blind user, meaning that the mental workload of the "sight" sample was higher than the one of "blind" sample.

The GSR ANOVA test also did not detect any impact from the methods or from the rounds, as it happened with the blind users in the Section \ref{subsubsec:results_gsr_temp_1}. Graphically the sight user variations were very similar in all methods. This is a different effect than the one observed in the blind users, which graphically showed different GSR distributions on different methods. The sight GSR also show a small variation between the rounds and methods, which means that the sight user were not stressed or had a low mental workload during the experiment.

Despite the proved and not proved tests, there is a consideration to be made. The sight sample group profiling. As already explained before, the profile of the “blind” sample group was very wide and that can impact negatively in their performance. But the opposite effect may had happened with the “sight” sample group. This group was composed basically by researchers and engineer students, people that are typically involved with computers and technological devices, aging from 22 to 31 with an average of 27.5 years. This may biased the results with better performance when using the HMD and being able to feel present inside a virtual environment.

Besides these results, the “sighted” sample also commented the experiment. They all felt a lot more insecurity when walking, exploring and even when hand guided by the researcher before the start of the round. The “blind” sample group was already used to bumping their body when exploring new closed quarters. The “sighted” group did not want that to happen and approached the furniture with a lot more caution. They also noticed the lack of precision of the haptic devices, but they did rely more on then to navigate.

%The processing of each data collected is rather similar and follows these steps:
%\begin{enumerate}
%    \item Separate the Blind sample and the Sight sample;
%    \item Check if the samples are normally distributed; \label{itm:results_shapiro} \\
%        If the data is normally distributed then it is possible to use other statistical analyses and verify the results statistically.
%    \item Check if the "blind" sample is statistically different then the "sight" sample; \label{itm:results_t_test} \\ 
%        This is one of the goals. To verify that the workload and the situation awareness of the blind participants are different from the sighted participants.
%    \item For the physiological data:
%    \begin{enumerate}
%        \item Calculate the variation between the scene and the baseline  in each method;
%        \item Calculate the variation between the scenes in each method;
%        \item Calculate the average variation in the group in each method.
%    \end{enumerate}
%    \item For the other data:
%    \begin{enumerate}
%        \item Calculate the variation between the scenes in each method; \label{itm:results_average_method_particpant}
%        \item Calculate the average variation in the group in each method.
%        \label{itm:results_average_method}
%    \end{enumerate}
%        The variation mentioned above is calculated as Equation \ref{eq:variation}:
%    \begin{equation}
%        \label{eq:variation}
%        Var_{ij} = \frac{Obs_j - Obs_i}{Obs_i}*100 [\%]
%    \end{equation}
%        Where:\\
%        $Var_{ij}$ = Variation between the two sequential observations \\
%        $Obs_i$ = First observation \\
%        $Obs_j$ = Second observation 
%    
%    \item Check if there are methods that are statistically different from the rest. \\
%        This is done by doing an analyses of variance (ANOVA) in the data. For this analyses is required that the observerd residues are normally distributed and the variance is constant. The residues are all plotted in the figures of the Appendix \ref{ap:figures}. The variance constancy can be verified by the box plot presented in the next sections.
%
%\end{enumerate}